---
title: "assig"
author: "Louis Duverger"
date: "27/11/2016"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)
```

## Loading and preprocessing the data

```{r}
library(dplyr)
library(xtable)
dt <- read.csv(file = "activity.csv", na.strings = "NA")
dt[,"date"]  <- as.Date(dt[,"date"], format = "%Y-%m-%d")
dt <- tbl_df(dt)
str(dt)
```

## Histogram of the total number of steps taken each day

```{r}
library(plotly)
dt1 <- dt %>% group_by(date) %>% summarize(total_steps = sum(steps))
plot_ly(data = dt1, x = dt1$total_steps, type='histogram')
```

## Mean and median number of steps taken each day
```{r}
dt2 <- dt %>% group_by(date) %>% summarize(total_steps = sum(steps))
dt2.mean <- mean(dt2$total_steps, na.rm = TRUE)
dt2.median <- median(dt2$total_steps, na.rm = TRUE)
```
The mean is `r dt2.mean` and the median is `r dt2.median`

## Time series plot of the average number of steps taken
```{r}
dt3 <- dt %>% group_by(date) %>% summarize('mean' = mean(steps))
plot_ly( data = dt3, x = dt3$date, y = dt3$mean, type = 'scatter', mode = 'lines')
```

## The 5-minute interval that, on average, contains the maximum number of steps

```{r}
dt4 <- dt %>% group_by(interval) %>% summarize('mean' = mean(steps, na.rm = TRUE))
interval.max <- dt4 %>% filter(dt4$mean == max(dt4$mean)) %>% select(interval)
```

The interval with the highest steps on average is `r interval.max`

## Code to describe and show a strategy for imputing missing data

We can use the method of the K nearest neighbours

## Code to describe and show a strategy for imputing missing data

```{r}
nas <- sum(is.na(dt$steps))
```
There are `r nas` in the data set.  The possible strategies are:
* substitute by the mean
* impute the value by looking at the neighbours (knn)

lets use a little bit of ML to impute the KNN. We wont bother doing CV to set the value of K.
```{r, cache=TRUE}
library(caret)
train <- dt %>% filter(!is.na(steps))
test <- dt %>% filter(is.na(steps))
knn <- train(data = train, steps ~., method ='knn')
test[,"steps"] <- predict(object = knn, newdata = test)
new.dt <- rbind(train,test )

dt5 <- new.dt %>% group_by(date) %>% summarize(total_steps = sum(steps))
plot_ly(data = dt5, x = dt5$total_steps, type='histogram')

dt5 <- dt %>% group_by(date) %>% summarize(total_steps = sum(steps))
dt5.mean <- mean(dt5$total_steps, na.rm = TRUE)
dt5.median <- median(dt5$total_steps, na.rm = TRUE)
```

## Histogram of the total number of steps taken each day after missing values are imputed

The mean is `r dt5.mean` and the median is `r dt5.median`
They are equal to the previous means and median, though the histogram is slightly different

## Panel plot comparing the average number of steps taken per 5-minute interval across weekdays and weekends
```{r}
library(lubridate)
library(ggplot2)
dt7 <- dt %>% mutate('wday'=wday(date)) %>% mutate('days' = ifelse(wday<6,'week','weekend'))
dt7 <- dt7 %>% group_by(days, interval) %>% summarize('mean_steps' = mean(steps, na.rm = TRUE))
g <- ggplot(data = dt7, aes(x = interval, y = mean_steps)) + geom_line() + facet_grid(days~.)
ggplotly(g)
```


